---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns

In SQL you can filter columns by using SELECT and rows by using WHERE.Let's look at the example using the Lahman baseball database. First, we should load the Lahmen package. We will also need the sqldf package.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Let's take a look at an example. Suppose we want to see the homerun totals for the 1927 Yankees. We could write:
```{r}
query<- "SELECT playerID,teamID,HR,yearID FROM Batting
WHERE teamID='NYA'and yearID=1927"
sqldf(query)
```
##ORDER BY

with SQL you can also order the rows in the output. For example, suppose you wanted to see every instance of a player hitting more than 50 homeruns, but you would like the players with the most homeruns to be at the top of this list. We could do this in the following way: 
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE HR>=50"

sqldf(query)
```
##Join
```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```
```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```
```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"

sqldf(query)
```

